### 2024-02-15 outputs/Table2_calculations/Table2_calculation.qmd
code for table 2

```{r}
#| echo: false
#| message: false
#| warning: false

# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ # 
######## B. META-ANALYSIS: for each of the hypotheses ######## 
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ #

# Data preparation: Split dataset in 3 based on outcome type
# 1) outgroup dataset
dat_outgroup <- dat[dat$PA_Category == 10 |
                      dat$PA_Category == 99, ]
length(unique(dat_outgroup$ID_R))#number of manuscripts (N_j)
# 2) conservative shift dataset
dat_conservatism <- dat[dat$PA_Category == 5 |
                          dat$PA_Category == 6 |
                          dat$PA_Category == 7 |
                          dat$PA_Category == 8 |
                          dat$PA_Category == 9 |
                          dat$PA_Category == 11, ]
length(unique(dat_conservatism$ID_R))#number of manuscripts (N_j)
# 3) rally-around-the-flag dataset
dat_rally <- dat[dat$PA_Category == 1 |
                   dat$PA_Category == 2 |
                   dat$PA_Category == 3 |
                   dat$PA_Category == 4, ]
length(unique(dat_rally$ID_R))#number of manuscripts (N_j)
# double check (total number of effect sizes; N_k)
count(dat_outgroup) + count(dat_conservatism) + count(dat_rally)

##### 1. Fit empty 3-level model using the meta3 function #####

### To replicate Table 2: call summary() function for each of outcome cluster
      ## coef(model...)[1] = Fishers' Z correlation coefficients (Zr)
### To replicate the correlation coefficients mentioned in the text, 
### call the FisherZinv() function for each outcome cluster.

# Outgroup hostility (Table 2)
set.seed(1234)
model_og_LBCI <- meta3L(y=Fisher, v=Variance_F, cluster=ID_R,
                       data = dat_outgroup, 
                       I2=c("I2q", "ICC"),
                       intervals.type = "LB",
                       model.name = "3 level outgroup LB")
summary(model_og_LBCI)#results presented in Table 2, as well as
#throughout the Supplementary Information (esp. Table B.3).
FisherZInv(coef(model_og_LBCI)[1])#rho reported in text

# Conservative shift (Table 2)
set.seed(1234)
model_cs_LBCI <- meta3L(y=Fisher, v=Variance_F, cluster=ID_R,
                       data = dat_conservatism, 
                       I2=c("I2q", "ICC"),
                       intervals.type = "LB",
                       model.name = "3 level conservatism LB")
summary(model_cs_LBCI)#results presented in Table 2, as well as
#throughout the Supplementary Information (esp. Table B.3).
FisherZInv(coef(model_cs_LBCI)[1])#rho reported in text

# Rally-around-the-flag (Table 2)
set.seed(1234)
model_rf_LBCI <- meta3L(y=Fisher, v=Variance_F, cluster=ID_R,
                       data = dat_rally, 
                       I2=c("I2q", "ICC"),
                       intervals.type = "LB",
                       model.name = "3 level rally LB")
summary(model_rf_LBCI)#results presented in Table 2, as well as
#throughout the Supplementary Information (esp. Table B.3).
FisherZInv(coef(model_rf_LBCI)[1])#rho reported in text

## Substantive interpretation --------------------------------------------------
## i.e., comparison to 2016 ANES Feeling Thermometers
anes2016 <- read_dta(file = "/cloud/project/raw/data/anes_timeseries_2016.dta")
#select variables to summarize
anes2016 <- anes2016 %>%
  select(immterm = V162313, 
         reptherm = V161096) 
#format(round(psych::describe(anes2016),3), nsmall=3): counts NA's
#recode NA's (!)
anes2016$immterm <- ifelse(anes2016$immterm %in% -9:-1, NA, 
                           identity(anes2016$immterm)) 
anes2016$reptherm <- ifelse(anes2016$reptherm %in% -99:-1, NA, 
                            identity(anes2016$reptherm))
#find correct SDs to use in calculations based on Paluck et al. (2021)
format(round(psych::describe(anes2016),3), nsmall=3) 
#the effect of terror on the feeling therm. toward immigrants. 6.9
cohens_d(r = FisherZInv(coef(model_og_LBCI)[1]))*27.285 
#the effect of terror on the feeling therm. toward Rep. party: 7.2
cohens_d(r = FisherZInv(coef(model_cs_LBCI)[1]))*27.330

## Practical implication -------------------------------------------------------
## i.e., what sample size is recommended based on this meta-analysis?
# Outgroup hostility
pwr.r.test(r = FisherZInv(coef(model_og_LBCI)[1]), sig.level = 0.05, power = 0.8) 
# Conservative shift
pwr.r.test(r = FisherZInv(coef(model_cs_LBCI)[1]), sig.level = 0.05, power = 0.8) 
# Rally-around-the-flag
pwr.r.test(r = FisherZInv(coef(model_rf_LBCI)[1]), sig.level = 0.05, power = 0.8)

```